---
title: "Installing CmdStan for EpiSewer"
output: github_document
---

This vignette describes how to install CmdStan on your system to use as a
backend for EpiSewer. If you cannot install CmdStan locally, you can
alternatively use the docker backend (see the
[docker vignette](https://adrian-lison.github.io/EpiSewer/articles/docker-backend.html)).

### Step 1: Install the EpiSewer package

Install the EpiSewer package from GitHub as usual:

```{r, eval = FALSE}
remotes::install_github("adrian-lison/EpiSewer", dependencies = TRUE)
```

### Step 2: Check the C++ toolchain

CmdStan requires a working C++ toolchain. You can check whether your system is
ready using:

```{r, eval = FALSE}
cmdstanr::check_cmdstan_toolchain()
```

If the toolchain is not set up, follow the instructions printed by the above
function. The [cmdstanr vignette](https://mc-stan.org/cmdstanr/articles/cmdstanr.html)
also contains platform-specific guidance for setting up the toolchain on
Windows, macOS, and Linux.

### Step 3: Install CmdStan

Once the toolchain is ready, install CmdStan using:

```{r, eval = FALSE}
cmdstanr::install_cmdstan(cores = 2) # use more cores to speed up installation
```

You can verify the installation was successful by checking the CmdStan version:

```{r, eval = FALSE}
cmdstanr::cmdstan_version()
```

### Step 4: Compile the EpiSewer models

The Stan models used by EpiSewer need to be compiled for your device. This is
only necessary once — after installing or updating the package — and can be
done using:

```{r, eval = FALSE}
EpiSewer::sewer_compile()
```

### Step 5: Load the package and run EpiSewer

```{r}
library(EpiSewer)
```

You can now run EpiSewer as described in the
[README](https://adrian-lison.github.io/EpiSewer/) and other vignettes. For
example, using the bundled example data:

```{r, eval = FALSE}
ww_result <- EpiSewer(
  data = ww_data_SARS_CoV_2_Zurich,
  assumptions = ww_assumptions_SARS_CoV_2_Zurich,
  fit_opts = set_fit_opts(
    sampler = sampler_stan_mcmc(
      iter_warmup = 500, iter_sampling = 500, chains = 4, seed = 42
      )
  )
)
```

```{r, eval = TRUE, include = FALSE}
ww_result <- ww_result_SARS_CoV_2_Zurich
```

### Step 6: Inspect results as usual

For example, plot the estimated reproduction number:
```{r R}
#| fig.alt: >
#|   Plot of the effective reproduction number over time.
plot_R(ww_result)
```

### Troubleshooting

If the models are not successfully compiled in step 4, first ensure that
CmdStan is properly set up by re-running step 2. You can also try updating
CmdStan to a newer version by re-running `cmdstanr::install_cmdstan()`.

If the problem persists, run `EpiSewer::sewer_compile(verbose = TRUE)` and
post the output in a new
[GitHub issue](https://github.com/adrian-lison/EpiSewer/issues), along with
your `cmdstanr::cmdstan_version()`.
