---
title: "Detailed example"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/example-",
  out.width = "100%",
  dpi = 330,
  message = FALSE,
  warning = FALSE
)
```

```{r, eval = TRUE, include = FALSE}
require(emo)
```

This vignette repeats the introductory example from the [README](../index.html), but shows how to specify the different modules and components in `EpiSewer` explicitly.

`r emo::ji("bulb")` For a conceptual overview over modules and components, we recommend to first read the `vignette("model-specification")` vignette.

### Loading the package

We assume that `EpiSewer` is already successfully installed (see [README](../index.html): installing the package).
```{r}
library(EpiSewer)
library(data.table)
library(ggplot2)
```

### Data

This is again the wastewater data from Zurich, Switzerland, provided by EAWAG (Swiss Federal Institute of Aquatic Science and Technology).
```{r}
data_zurich <- SARS_CoV_2_Zurich
```

As in the introductory example, we keep only measurements that were made on Mondays and Thursdays to test `EpiSewer` on sparse data.
```{r}
measurements_sparse <- data_zurich$measurements[,weekday := weekdays(data_zurich$measurements$date)][weekday %in% c("Monday","Thursday"),]
head(measurements_sparse, 10)
```

### Modeling

In the [README](../index.html), we used a shortcut to collect all observation data via `sewer_data()` and all assumptions via `sewer_assumptions()`. These are convenience functions to make the `EpiSewer` command more concise. In this example however, we will explicitly supply the data and assumptions to the relevant modules and components (see the `vignette("model-specification")` vignette).

#### Measurements

We start with the first module, the observed measurements. The module can be specified as follows:
```{r}
ww_measurements <- model_measurements(
  concentrations = concentrations_observe(measurements = measurements_sparse),
  noise = noise_estimate(),
  LOD = LOD_none()
)
```

As we see, the module is defined using the `model_measurements()` function. The arguments of this function represent the different module components. In this case, there is:

- a `concentrations` component: defines pathogen concentrations in the wastewater
- a `noise` component: defines measurement noise
- a `LOD` component: defines a limit of detection in wastewater samples

We used a specific modeling function for each of the components. The suffix of each function explains what we modeled:

- `concentrations_observe`: The concentrations are modeled as observation data to which the model can be fitted.
- `noise_estimate`: The degree of measurement noise is estimated from the data.
- `LOD_none`: We do not model a limit of detection. This also means that zero measurements are dropped.

We can also print the module object to see what we have modeled:

```{r}
ww_measurements
```


`r emo::ji("exclamation_mark")` Note that the modeling functions above have further arguments, and `EpiSewer` will use default settings if they are not specified. For example, we could set a custom prior on the measurement noise as follows:

```{r,}
ww_measurements2 <- model_measurements(
  concentrations = concentrations_observe(measurements = measurements_sparse),
  noise = noise_estimate(cv_prior_mu = 0.5, cv_prior_sigma = 0.1), # prior on coefficient of variation
  LOD = LOD_none()
)
```

`r emo::ji("right_arrow")` To find out which options are available for a given modeling function, check out its function documentation.

Before we move to the next module, remember that there can be multiple modeling options for a component, and `EpiSewer` offers one modeling function for each option. To see all available options, you can use `component_functions()`:

```{r}
component_functions("LOD")
```

For example, we could assume a limit of detection of `2.56 gc/mL` (based on a lab experiment by EAWAG, this was the lowest concentration at which still 95% of replicates were positive) as follows:
```{r,}
ww_measurements3 <- model_measurements(
  concentrations = concentrations_observe(measurements = measurements_sparse),
  noise = noise_estimate(cv_prior_mu = 0.5, cv_prior_sigma = 0.1), # prior on coefficient of variation
  LOD = LOD_assume(limit = 2.56, prob = 0.95)
)
```

If our measurements are based on digital PCR (dPCR), we can also model the LOD
with an explicit model of dPCR noise. Note that we also have to adapt the noise component accordingly.

```{r,}
ww_measurements4 <- model_measurements(
  concentrations = concentrations_observe(measurements = measurements_sparse),
  noise = noise_estimate_dPCR(), # also model dPCR noise
  LOD = LOD_estimate_dPCR()
)
```

#### Sampling

Next is the `sampling` module. It models the process of sampling from the wastewater:

```{r}
ww_sampling <- model_sampling(
  outliers = outliers_none(),
  sample_effects = sample_effects_none()
)
```

The sampling module currently has two component. The outlier component allows to automatically detect outlier observations during model fitting, making transmission dynamic estimates more robust. The sample effects component allows to model the effects of sample characteristics (like time from sampling until processing, which can lead to degradation) on the detectable pathogen concentrations. In this simple example, we do not model outliers or sampling effects.

#### Sewage

The `sewage` module describes aspects of the sewage system:

```{r}
ww_sewage <- model_sewage(
  flows = flows_observe(flows = data_zurich$flows),
  residence_dist = residence_dist_assume(residence_dist = c(1))
)
```

It includes two components:

- `flows`: Daily flow volumes at the sampling site. Here we use the measured flows at the treatment plant in Zurich, specified using `flows_observe()`.
- `residence_dist`: Sewer residence time distribution for pathogen particles. In large sewage systems, particles may travel longer than a day depending on where and when they were shed into the wastewater. This component allows to model the corresponding delay distribution. We here use the modeling function `residence_dist_assume()` to indicate that we make an assumption about the residence time. In our case, we assume that particles arrive at the sampling site within the same day (indicated by the vector `c(1)`, which puts 100% probability on a residence time of 0 days).


#### Shedding

The `shedding` module describes how infected individuals shed pathogen particles into the wastewater:

```{r}
ww_shedding <- model_shedding(
  shedding_dist = shedding_dist_assume(
    get_discrete_gamma(gamma_shape = 0.929639, gamma_scale = 7.241397), shedding_reference = "symptom_onset"),
  incubation_dist = incubation_dist_assume(get_discrete_gamma(gamma_shape = 8.5, gamma_scale = 0.4)),
  load_per_case = load_per_case_calibrate(cases = data_zurich$cases),
  load_variation = load_variation_estimate()
)
```


The shedding module requires a number of assumptions (see explanations in the [README](../index.html)):

- `shedding_dist` and `incubation_dist`: We use the same shedding load distribution (by days since symptom onset) as before. We also need to assume an incubation period distribution.
- `load_per_case`: The load per case assumption is calibrated based on observed case data. In the [README](../index.html), we supplied the case data via `sewer_data()`. Here, we now supply it directly to `load_per_case_calibrate()`.
- `load_variation`: `EpiSewer` offers the option to model individual-level variation in shedding, and this is recommended as a default.

#### Infections

The `infections` module describes the underlying process leading to infected individuals:

```{r}
ww_infections <- model_infections(
  generation_dist = generation_dist_assume(get_discrete_gamma_shifted(gamma_mean = 3, gamma_sd = 2.4)),
  R = R_estimate_splines(),
  seeding = seeding_estimate_rw(),
  infection_noise = infection_noise_estimate()
)
```

The module has the following epidemiological components:

- `generation_dist`: We use the same generation time distribution as before.
- `R`: This the effective reproduction number, our main parameter of interest. We here estimate it via smoothing B-splines (`R_estimate_splines()`), but there are other smoothing approaches available (see `component_functions("R")`).
- `seeding`: The renewal model used by `EpiSewer` requires a seeding phase during which the reproduction number cannot be modeled. We thus let the initial infections be estimated using a simple geometric random walk model.
- `infection_noise`: We also model noise in the infection process, meaning that `EpiSewer` will fit a stochastic infection model. This makes sense in most cases.

#### Forecast

Last but not least, there is the `forecast` module. This module describes how the generative model defined by the other modules is used to produce forecasts of Rt, infections, concentrations and other quantities. By default, as in our [README](../index.html) example, we produce no forecasts by effectively setting the forecast horizon to zero.
```{r}
ww_forecast <- model_forecast(
  horizon = horizon_none()
)
```

If we want to produce forecasts, we can define the forecast horizon (e.g. 7 days) as follows:

```{r}
ww_forecast2 <- model_forecast(
  horizon = horizon_assume(horizon = 7)
)
```

The forecast module currently only has a single component to define the forecast horizon. In the future, this may be extended to allow users to customize how forecasts are obtained.

#### Side notes

We have now specified all six `EpiSewer` modules. While this seems like a lot of code, note that the majority of component specifications are optional and have sensible defaults. In practice, you only need to explicitly write out components that need certain assumptions or that you want to customize. In the present example, we also (unnecessarily) specified the `EpiSewer` defaults. For example, the `infections` module described above could simply be written as

```{r}
ww_infections <- model_infections(
  generation_dist = generation_dist_assume(get_discrete_gamma_shifted(gamma_mean = 3, gamma_sd = 2.4))
)
```

because the `R`, `seeding`, and `infection_noise` component used the defaults. No default is given for the generation time distribution, since it is pathogen-specific.

`r emo::ji("bulb")` An advantage of coding each component explicitly is that your modeling decisions are clearly documented.

### Settings

Before starting the estimation, let's have a quick look at some settings for model fitting and post-processing.

The `fit_opts` settings describe how the model is fitted. The first argument, `model` tells EpiSewer where to find the `stan` files with the model code. The default is the `EpiSewer` package itself, but you can also tweak this setting to provide your own customized model files (see `model_stan_opts()` for details).
The `sampler` argument specifies the sampling algorithm used for model fitting. Currently, only sampling via stan's MCMC algorithm is supported. The function `sampler_stan_mcmc()` allows you to pass various arguments to stan, including number of chains, iterations, and much more.
```{r}
ww_fit_opts <- set_fit_opts(
  model = model_stan_opts(package = "EpiSewer"),
  sampler = sampler_stan_mcmc(
    iter_warmup = 500,
    iter_sampling = 500,
    chains = 4,
    parallel_chains = 4, # run all chains in parallel
    seed = 42
  )
)
```

The `results_opts` argument specifies what results to return after model
fitting. We specify `fitted=TRUE`, which means that the whole fitted model
object will be stored. Setting this to `fitted=FALSE` saves memory  - but certain
things like detailed diagnostics or extracting draws of additional parameters
will not be possible. We also specify that parameters of interest should be
summarized with 50% and 95% credible intervals (in addition to the mean and
median). Moreover, we want to explicitly store 50 posterior samples for the Rt
and infections time series (useful for spaghetti plots etc.).

```{r}
ww_results_opts <- set_results_opts(
  fitted = TRUE,
  summary_intervals = c(0.5, 0.95),
  samples_ndraws = 50
)
```

### Estimation

Now we have everything we need. To fit the model, we simply pass all the
EpiSewer modules and settings defined above to the `EpiSewer()` function.

```{r, eval = FALSE}
ww_result <- EpiSewer(
  measurements = ww_measurements,
  sampling = ww_sampling,
  sewage = ww_sewage,
  shedding = ww_shedding,
  infections = ww_infections,
  forecast = ww_forecast,
  fit_opts = ww_fit_opts,
  results_opts = ww_results_opts
)
```

Note that this time we don't need a `data` or `assumptions` argument, because we provided all data and assumptions to the respective module components.

The above model is identical to the default model fitted in the [README](../index.html), so we don't repeat the results again.

```{r, eval = TRUE, include = FALSE}
ww_result <- EpiSewer(
  measurements = ww_measurements,
  sampling = ww_sampling,
  sewage = ww_sewage,
  shedding = ww_shedding,
  infections = ww_infections,
  forecast = ww_forecast,
  fit_opts = ww_fit_opts,
  results_opts = ww_results_opts,
  run_fit = FALSE
)
```

Remember that you can print a summary of the full modeling details from the job object:
```{r}
ww_result$job$model
```

### Full model

For your convenience, we here provide the full model using the explicit component specifications in one code block. This can serve as a starting point to adapt the specification and try out different modeling options (check out the function documentation for the module functions, or try out the `component_functions()` helper).

Also, don't forget that the modeling functions have further arguments to customize priors and other modeling details. As always, the function documentation is your friend! 

```{r, eval = FALSE}
measurements_sparse <- data_zurich$measurements[
  , weekday := weekdays(data_zurich$measurements$date)
  ][weekday %in% c("Monday","Thursday"),]

ww_measurements <- model_measurements(
  concentrations = concentrations_observe(measurements = measurements_sparse),
  noise = noise_estimate(),
  LOD = LOD_none()
)

ww_sampling <- model_sampling(
  outliers = outliers_none(),
  sample_effects = sample_effects_none()
)

ww_sewage <- model_sewage(
  flows = flows_observe(flows = data_zurich$flows),
  residence_dist = residence_dist_assume(residence_dist = c(1))
)

ww_shedding <- model_shedding(
  shedding_dist = shedding_dist_assume(get_discrete_gamma(gamma_shape = 0.929639, gamma_scale = 7.241397), shedding_reference = "symptom_onset"),
  incubation_dist = incubation_dist_assume(get_discrete_gamma(gamma_shape = 8.5, gamma_scale = 0.4)),
  load_per_case = load_per_case_calibrate(cases = data_zurich$cases),
  load_variation = load_variation_estimate()
)

ww_infections <- model_infections(
  generation_dist = generation_dist_assume(get_discrete_gamma_shifted(gamma_mean = 3, gamma_sd = 2.4)),
  R = R_estimate_splines(),
  seeding = seeding_estimate_rw(),
  infection_noise = infection_noise_estimate()
)

ww_forecast <- model_forecast(
  horizon = horizon_none()
)

ww_fit_opts <- set_fit_opts(
  model = model_stan_opts(package = "EpiSewer"),
  sampler = sampler_stan_mcmc(
    iter_warmup = 500,
    iter_sampling = 500,
    chains = 4,
    parallel_chains = 4, # run all chains in parallel
    seed = 42
  )
)

ww_results_opts <- set_results_opts(
  fitted = TRUE,
  summary_intervals = c(0.5, 0.95),
  samples_ndraws = 50
)

ww_result <- EpiSewer(
  measurements = ww_measurements,
  sampling = ww_sampling,
  sewage = ww_sewage,
  shedding = ww_shedding,
  infections = ww_infections,
  forecast = ww_forecast,
  fit_opts = ww_fit_opts,
  results_opts = ww_results_opts
)
```

