#### Shedding load per case
In addition to the shedding load distribution, we also require a scaling factor 
that describes how much total viral material is shed by the average infected 
individual and is detectable at the sampling site. We call this scaling factor
*shedding load per case* because it scales the estimated number of cases/infections
in our model. The shedding load per case (gc/person) depends both on biological factors as 
well as on the sewage system.

Fortunately, if we are only interested in the effective reproduction number, 
a rough estimate for the shedding load per case is sufficient: 
getting it wrong cannot strongly bias the reproduction number estimates, and 
will distort the uncertainty of estimates only in rather extreme cases.

If we also have case data available (ideally catchment-specific, or we might
down-scale case counts from a higher aggregation level to roughly the size
of our catchment), then we can use the function `suggest_load_per_case` to 
obtain a rough estimate of a suitable shedding load per case. The argument 
`ascertainment_prop` can be used to account for underdetection of infections. 
For simplicity, we here assume `ascertainment_prop=1`, meaning that 100% of 
infections become confirmed cases (which is often not realistic).
```{r}
suggest_load_per_case(
  data_zurich$measurements,
  data_zurich$cases,
  data_zurich$flows,
  ascertainment_prop = 1
)
```

Hence we will assume `1e+11 gc/person` as the average shedding load.
```{r}
load_per_case <- 1e+11
```
